---
title: "Dans data script"
format: html
editor: visual
---

## Data analyze for Dan's data with LMER model

## Main analyze file! 

1 Load libraries

```{r}
library(conflicted)  
library(tidyverse)
library(lmerTest)
library(sjPlot)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
conflicts_prefer(lmerTest::lmer)
```

2 Data from CSV-file to the structure

```{r}
raw <- here::here("data/raw_data.csv") |> readr::read_csv()
```

3 If you need to check all the raw data, you can do it here:

```{r}
view(raw)
```

4 Next we select only certain columns. Change last value (for example: '**SF12MCS**') when you work with different predictor variable.

```{r}
df <- raw |> 
  select(ParticipantCode, Group, Time, BREF_PSYCHOL)
```

5 View data to check that everything is in order. Good idea to check against excel-file. We rename Outcome value as outcome so we only have to change it to one place (number 4 above).

```{r}
colnames(df) <- c("ParticipantCode", "Group", "Time", "Outcome")
view(df)
#str(df)
```

6 Fit linear mixed effects model. We call this model simply 'M1'. Run summary to see that values look right. Change value '**SF12MCS**' when you work with different column.

```{r}
M1 <- lmer(Outcome ~ Time*Group+(1|ParticipantCode),data=df)
summary(M1)
```

7 Tab model:

```{r}
tab_model(M1)
```
