---
title: "Descriptive Statistics"
author: "Petri Haavisto"
format: html
editor: visual
---

## Descriptive statistics to summarize the characteristics of the data

1 First load libraries:

```{r}
library(conflicted)  
library(tidyverse)
library(sjPlot)
library(lmerTest)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
conflicts_prefer(lmerTest::lmer)
```

2 Next we read raw_data.csv into a tibble. Tibble is a modern version of data frame

```{r}
raw <- here::here("data/raw_data.csv") |> readr::read_csv() 
```

3 Next we filter away unnecessary columns and leave only data of interests.

```{r}
(df <- raw |> 
  select(ParticipantCode, Group, Gender_1M_2F))
```

4 Rename Columns.

```{r}
colnames(df) <- c("participant", "group", "gender")
```

5 Remove doubles (we need only one row per participant).

```{r}
df <- df |> distinct(participant, group, gender)
```

6 We can have a look of structure we filtered and make sure everything is all right.

```{r}
view(df)
```

7 Create a gender plot by group.

```{r}
df |> 
  ggplot(aes(group, fill=gender))+
  geom_bar(position = "dodge", alpha = 0.5)+ #dodge separates bars
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())+
  labs(title = "Gender by group",
       x = "Group",
       y = "n")
  ggsave(filename = "../images/gender_by_group.png")
```

8 Also a gender plot but now with the facet.

```{r}
df |> 
  ggplot(aes(x = gender))+
  geom_bar(aes(fill=gender), alpha = 0.5)+ 
  facet_wrap(~ group)+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())+
  labs(title = "Gender by group",
       x = "",
       y = "n")
  ggsave(filename = "../images/gender_by_group_facet.png")
```

9 Next same but with education

```{r}
(df2 <- raw |> 
  select(ParticipantCode, Group, EDU_1Primary_2Seconday_3Tertiary))
```

10 Rename columns

```{r}
colnames(df2) <- c("participant", "group", "education")
```

11 Remove doubles (we need only one row per participant.

```{r}
df2 <- df2 |> distinct(participant, group, education)
```

12 We can have a look of structure we filtered and make sure everything is all right.

```{r}
view(df2)
```
