---
title: "visualizing data"
author: "Petri Haavisto"
format: html
editor: visual
---

## Preparations before we start

Installing tidyverse and "here" packages (one time\*). Here function returns the path of your current project. These are installed from CRAN (Comprehensive R Archive Network).

\*unfortunately packages need to be re-installed when R/RStudio is re-installed/updated

```{r}
# Install from CRAN
install.packages("tidyverse")
install.packages("here")
```

Let's load whole tidyverse ecosystem

```{r}
library(tidyverse)
```

## Start working with data

We start by loading the data file

```{r}
hma_tib <- here::here("data/raw_data.csv") |> readr::read_csv()
```

Let's have glimpse of data

```{r}
glimpse(hma_tib)
```

That wasn't very helpful so let's divide data to different groups starting with first one:

```{r}
inter1 <- hma_tib |> 
  filter(Group == "InterV_PRE_1")
```

glimpse of new group:

```{r}
glimpse(inter1)
```

Plot of relationships of InterV-group

```{r}
#na.rm removes empty values without warning
#added fourth group here (4 = "Unknown")

gb <- ggplot(inter1, aes(x  = Living_with_1Single_2Married_3living_together)) +
  geom_bar() +
  theme(axis.title.x = element_blank())
gb +  scale_x_discrete(limits=c("Single", "Married", "Living together", "Unknown"))
```
