---
title: "Muay Thai t-test"
author: "Petri Haavisto"
format: html
editor: visual
---

## Muay Thai Group *t*-test

1\) Librariest to use

```{r}
library(conflicted)
library(tidyverse)
library(pastecs)
library(reshape2)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
```

2\) Read raw data to a tibble

```{r}
raw <- here::here("data/raw_muaythai_data.csv") |> readr::read_csv()
```

3\) Next we filter only columns: ParticipantCode, Time and OUTCOME-variable

```{r}
(df <- raw |> 
  select(ParticipantCode, Time, PERMA_WellBeing))
```

4\) Rename columns

```{r}
colnames(df) <- c("id", "Time", "Outcome")
#view(df)
```

5\) Create wide view

```{r}
intervention_group <- dcast(df, id~Time, value="Outcome")
```

6\) Rename columns and CHECK THE DATA

```{r}
colnames(intervention_group) <- c("id", "Pre", "Post")
view(intervention_group)
```

7.1) After this, we don't need id column any more

```{r}
intervention_group <- intervention_group |>  select(-id)
```

7.2) Descriptive statistics

```{r}
stat.desc(intervention_group, basic = FALSE, norm = FALSE)

```

8\) Calculate differences

```{r}
intervention_group$diff <- intervention_group$Post - intervention_group$Pre
mean(intervention_group$diff)
view(intervention_group)
```

9\) We can do normality tests for differences using the stat.desc() function.

```{r}
norm_test <- stat.desc(intervention_group$diff, basic = FALSE, desc = FALSE, norm = TRUE)
print (norm_test[6])
```

10\) Do the t-test and view results

```{r}
dep_t_test <- t.test(intervention_group$Post, intervention_group$Pre, paired = TRUE)
dep_t_test
```

11\) Calculate the effect size. 0.5 is threshold for a large effect.

```{r}
t <- dep_t_test$statistic[[1]]
df <- dep_t_test$parameter[[1]]
r <- sqrt(t^2/(t^2+df))
round(r, 3)
```
