---
title: "Martial Arts raw results"
author: "Petri Haavisto"
format: html
editor: visual
---

## Adjusted values for Error Bar

Pipe=Ctrl+Shift+M.

Left Arrow = Alt and - (minus)

1\) First libraries we are going to use:

```{r}
library(conflicted)  
library(tidyverse)
library(sjPlot)
library(reshape2)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
```

2\) Next we read raw_data.csv into a tibble. Tibble is a modern version of data frame.

```{r}
raw <- here::here("data/raw_muaythai_data.csv") |> readr::read_csv()
```

3\) Next we filter only certain columns: ParticipantCode, Group, Time and OUTCOME-variable

```{r}
(df <- raw |> 
  select(ParticipantCode,Time, SF12MCS))
```

4\) Rename OUTCOME variable as outcome, so we don't have to rename it to all possible places. Also shorten the Participantcode to id.

```{r}
colnames(df) <- c("id", "Time", "Outcome")
```

6\) We can have a look of structure we filtered and make sure everything is all right.

```{r}
view(df)
```

7\) Create wide view.

```{r}
wide_df <- dcast(df, id~Time, value="Outcome")
```

9\) Rename columns

```{r}
colnames(wide_df) <- c("id", "Pre", "Post")
view(wide_df)
```

------------------------------------------------------------------------

### Calculations starts here

1\) Calculate pMean for every participant

```{r}
wide_df$pMean <- (wide_df$Pre + wide_df$Post)/2 
```

2\) Calculate Grand Mean

```{r}
grandMean <- mean(c(wide_df$Pre, wide_df$Post))
grandMean
```

3\) Calculate Adjustment factor (grandMean - pMean)

```{r}
wide_df$adj <- grandMean - wide_df$pMean
```

4\) Create adjusted values for pre and post

```{r}
wide_df$pre_adj <- wide_df$Pre + wide_df$adj
wide_df$post_adj <- wide_df$Post + wide_df$adj
view(wide_df)
```

5\) we can put data to long form again

```{r}
adjusted_df <- melt(wide_df, id = c("id", "Pre", "Post", "pMean", "adj"), measured = c("pre_adj", "real_adj"))
view(adjusted_df)
```

6\) Remove unwanted columns

```{r}
adjusted_df <- adjusted_df[, -c(2:5)]
names(adjusted_df) <- c("id", "Time", "Variable")
view(adjusted_df)
```

7\) Let's graph error bars based on adjusted data

```{r}
bar <- ggplot(adjusted_df, aes(Time, Variable))
```
