---
title: "SF12MCS"
author: "Petri Haavisto"
format: html
editor: visual
---

## SF12MCS

Libraries:

```{r}
library(conflicted)  
library(tidyverse)
library(sjPlot)
library(lmerTest)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
conflicts_prefer(lmerTest::lmer)
```

Next raw data to the structure:

```{r}
raw <- here::here("data/raw_data.csv") |> readr::read_csv()
```

Next we filter only certain columns: ParticipantCode, Group, Time and SF12MCS. (Pipe=Ctrl+Shift+M)

```{r}
df <- raw |> 
  select(ParticipantCode, Group, Time, SF12MCS)
```

View data to check that everything is in order.

```{r}
view(df)
```

Linear mixed effects model:

```{r}
M1 <- lmer(SF12MCS ~ Time*Group+(1|ParticipantCode), REML=F, data=df)
summary(M1)
```

Tab model:

```{r}
tab_model(M1)
```

Fit another linear mixed-effects model:

```{r}
M2 <- lmer(SF12MCS ~ Time+Group+Time*Group+(1|ParticipantCode), REML = F, data=df)
summary(M2)
```

Comparison of M1 and M2 with anova tells that there is no difference with these two models so let's use M1:

```{r}
anova(M1, M2)
```

Let's run ANOVA analysis to the model and calculate effect sizes.

```{r}
avc = anova(M1)
avc <- avc[,c(1:3,5,6)]
colnames(avc) <- c("sum sq", "mean sq", "DF", "F-value", "p-value")
view(avc)
```
