---
title: "Testing lmer with sleepstudy"
author: "Petri Haavisto"
format: html
editor: visual
---

## Testing Mixed models with sleep study data

```{r}
library(conflicted)  
library(tidyverse)
library(sjPlot)
library(lmerTest)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
conflicts_prefer(lmerTest::lmer)
```

Comparison of different models: (**y \~ x means "y predicted by x")**

lm(y \~ x) \## *linear model*

glm(y \~ x, family = binomial) \## *generalized linear model*

lmer(y \~ x, + (1 \| group)) \## *multilevel models*

glmer(y\~x + (1 \| group), family = binomial) \## *multilevel logistic or Poisson regression*

```{r}
d <- here::here("data/sleepstudy.csv") |> readr::read_csv()
```

```{r}
view(d)
cols = rainbow(5) #make colors for 5 subjects
plot(d$Days, d$Reaction, col=cols[d$Subject], pch=16)
```

See what happens if we fit a regular linear regression and plot the regression line.

```{r}
m = lm(Reaction ~ Days, data=d)
tab_model(m)
```

And we can visualize this.

```{r}
plot(d$Days, d$Reaction, col=cols[d$Subject], pch = 16)
abline(coef(m)[1], coef(m)[2])
```

Next we improve this with Multilevel model with random intercepts. In other words we add random component of another level of data.

```{r}
m_ri = lmer(Reaction ~ Days + (1 | Subject), data=d)
```

And we can plot it.

```{r}
plot(d$Days, d$Reaction, col = cols[d$Subject], pch = 16)
for (i in 1:5){ ##for each Subject
  abline(coef(m_ri)$Subject[i,1], coef(m_ri)$Subject[i,2], col = cols[i])
}

```

Let's have a look of the model

```{r}
tab_model(m_ri)
```

Next we can expand our multilevel model by using random intercepts AND **random slopes**.

```{r}
#random intercept and random slope(Days)
m_rs = lmer(Reaction ~ Days + (1 + Days | Subject), data=d)
tab_model(m_rs)
```

And then plot it.

```{r}
plot(d$Days, d$Reaction, col = cols[d$Subject], pch = 16)
for (i in 1:5){ ##for each Subject
  abline(coef(m_rs)$Subject[i,1], coef(m_rs)$Subject[i,2], col = cols[i])
}
```

## Comparing multilevel models

By comparing models, we can see how the variance at different levels changes. Here we first make a base model, which is a random intercepts model without any predictor variables. Then we add Days at the individual level, and finally we add the random slopes for Days.

```{r}
m_base = lmer(Reaction ~ (1 | Subject), data = d)
m_1 = lmer(Reaction ~ Days + (1 | Subject), data = d)
m_2 = lmer(Reaction ~ Days + (1 + Days | Subject), data = d)
anova(m_base, m_1, m_2)
```

```{r}
tab_model(m_base, m_1, m_2)
```
