---
title: "Spider t-test"
author: "Petri Haavisto"
format: html
editor: visual
---

## Spider dependent t-test

1\) Load libraries

```{r}
library(conflicted)  
library(tidyverse)
library(pastecs)
```

2\) Create test data frame

```{r}
picture <- c(30,35,45,40,50,35,55,25,30,45,40,50)
real <- c(40,35,50,55,65,55,50,35,30,50,60,39)
spider <- data.frame(picture, real)
```

3)  Descriptive statistics

```{r}
stat.desc(spider, basic = FALSE, norm = TRUE)
```

4\) Using the data compute differences between picture and real conditions and check the normality for these differences.

```{r}
spider$diff <- spider$real - spider$picture
view(spider)
```

5\) We can do normality tests using the stat.desc() function

```{r}
stat.desc(spider$diff, basic = FALSE, desc = FALSE, norm = TRUE)
```

6\) Do the t-test and view results

```{r}
dep_t_test <- t.test(spider$real, spider$picture, paired = TRUE)
dep_t_test
```

7\) Calculate the effect size

```{r}
t <- dep_t_test$statistic[[1]]
df <- dep_t_test$parameter[[1]]
r <- sqrt(t^2/(t^2+df))
round(r, 3)
```

8\) Let's plot box plot and error bars showing confidence intervals for spider data. First data to the long format.

```{r}
Group <- gl(2, 12, labels= c("Picture", "Real Spider"))
Anxiety <- c(30, 35, 45, 40, 50, 35, 55, 25, 30, 45, 40, 50, 40, 35, 50, 55, 65, 55, 50, 35, 30, 50, 60, 39)
spiderLong <- data.frame(Group, Anxiety)
view(spiderLong)
```

9\) Boxplot

```{r}
spiderBoxplot <- ggplot(spiderLong, aes(Group, Anxiety))
spiderBoxplot + geom_boxplot() + labs(x = "Type of Stimulus", y = "Anxiety") +
scale_y_continuous(limits = c(0, 100), breaks = seq(from = 0, to = 100, by = 10))
```

10\) Error bar chart:

```{r}
bar <- ggplot(spiderLong, aes(Group, Anxiety))
bar + stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") +
stat_summary(fun.data = mean_cl_normal, geom = "pointrange") + labs(x = "Type of
Stimulus", y = "Anxiety") + scale_y_continuous(limits = c(0, 65), breaks = seq(from= 0, to = 65, by = 10))
```
