---
?---
title: "visualizing data"
author: "Petri Haavisto"
format: html
editor: visual
---

## Preparations before we start

Installing tidyverse and "here" packages (one time\*). Here function returns the path of your current project. These are installed from CRAN (Comprehensive R Archive Network).

\*unfortunately packages need to be re-installed when R/RStudio is re-installed/updated

```{r}
# Install from CRAN
install.packages("tidyverse")
install.packages("here")
```

```{r}
install.packages("lmerTest")
```

Let's load whole tidyverse ecosystem. "Conflicted" is way to tackle possible conflict between packages.

```{r}
library(conflicted)  
library(tidyverse)
library(lmerTest)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
```

We start by loading the data file

```{r}
hin
```

Let's have glimpse of data

```{r}
glimpse(hma_tib)
```

That wasn't very helpful so let's divide data to different groups starting with InterV_PRE_1 :

```{r}
inter1 <- hma_tib |> 
  filter(Group == "InterV_PRE_1")
```

glimpse of new group:

```{r}
glimpse(inter1)
```

Plot of relationships of InterV_PRE_1 -group

```{r}
#na.rm removes empty values without warning
#added fourth group here (4 = "Unknown")

gb <- ggplot(inter1, aes(x  = Living_with_1Single_2Married_3living_together)) +
  geom_bar() +
  theme(axis.title.x = element_blank())
gb +  scale_x_discrete(limits=c("Partner", "Single", "Married", "Unknown"))

#Once you have created a plot, you can save it to projects images folder
ggsave(filename = "relationship_group1.png", path = here::here("images"))
```

Next let's select only certain information of first group: code, group, gender, relationship and SF12MCS.

```{r}
inter1_lite <- inter1 |> 
  select(ParticipantCode, Group, Gender_1M_2F, Living_with_1Single_2Married_3living_together, SF12PCS, SF12MCS )
```

```{r}
inter1_lite
```

```{r}
  ggplot(interlaes(SF12PCS, SF12MCS)) +
  geom_point()
```

Let's calculate SF12MSC mean of group 1:

```{r}
# Convert the character data type of SF12MSC to numeric
inter_lite1$SF12MCS <- as.numeric(inter_lite1$SF12MCS)
# Calculate the average SF12MSC score
mean_score <- mean(inter_lite1$SF12MCS, na.rm = TRUE)
message <- paste("The mean Score is: ",mean_score)
print(message)
```
