---
title: "weight_subjects"
author: "Petri Haavisto"
format: html
editor: visual
---

## Linear mixed effect models - random slopes and interactions

```{r}
library(conflicted)  
library(tidyverse)
library(sjPlot)
library(lmerTest)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
conflicts_prefer(lmerTest::lmer)
```

import the csv file

```{r}
df <- here::here("data/weight_subject.csv") |> readr::read_csv()
```

We can have look of data.

```{r}
view(df)
```

Fitting a model. (arrow with ALT + - )

```{r}
M4 <- lmer(Weight ~ Weeks+(1|Subjects), REML=F, data=df)
summary(M4)
```

Above Intercept 89.7750 and Weeks(slope) -2.9750 are fixed effects.

Extract random effects. Here for example first person intercept is 11.391193 above the overall intercept. In this case his starting weight is 89.7750 + 11.391193 = 101.166...

```{r}
ranef(M4)
```

Next let's allow also random slopes since people lose weight with different speeds. So we added Weeks inside the parenthesis.

```{r}
M3 <- lmer(Weight ~ Weeks+(1 + Weeks|Subjects), REML=F, data=df)
summary(M3)
```

Let's have look. Now Subjects have different slopes as well.

```{r}
ranef(M3)
```

Above we can see for example that subject 1 has weeks slope -0.3533220 steeper than average slope. From **Corr -0.88** we can see that people with higher initial weight lose weight faster than people with lower initial weight (this is expected).

Next we add Diet to fixed effects.

```{r}
M2 <- lmer(Weight ~ Weeks+Diet+(1|Subjects), REML=F, data=df)
summary(M2)
```

Above we can see that individual on Diet B has about 16 kg less weight at the starting point.

Let's have look:

```{r}
ranef(M2)
```

Next we allow different Diets to have different slopes.

```{r}
M1 <- lmer(Weight ~ Weeks*Diet+(1|Subjects), REML=F, data=df)
summary(M1)
```

Above we can see that slope on Diet A (that is our baseline) is -3.4000 and slope on Diet B is -3.4 + 0.85 = -2.55. Notice that interaction 0.06630 tells us that there is no significant difference in slopes of these two Diets so there is no significant difference in weight loss between diets.

Let's do **ANOVA**.

```{r}
anova(M1, M2)
```

Above Pr(\>Chisq) value 0.06092 tells us that M1 is not significantly better model than M2.
